### 2024-05-14 thesis/appendix.tex
correct names in contents,list of fig and tables
May 13th 2024

\chapter{Appendix}
\section{Feature Map Normalization}\label{appendix:new_feature}
\begin{proof}
\begin{align*}
    \|
    \phi^{new}(x)\|_{\mathcal{H}}^{2} &= \left\|\frac{\phi(x)}{\|\phi(x)\|_{\mathcal{H}}}
    \right\|_{\mathcal{H}}^{2}
    \\
    &=
    \left\|
    \frac{\phi(x)}
    {\sqrt{k(x,x)}}
    \right\|_{\mathcal{H}}^{2}
    \\
    &=
    \langle
    \frac{\phi(x)}
    {\sqrt{k(x,x)}}
    ,
    \frac{\phi(x)}
    {\sqrt{k(x,x)}}
    \rangle_{\mathcal{H}}
    \\
    &=
    \frac{1}{\sqrt{k(x,x)^{2}}}
    \langle
    \phi(x)
    ,
    \phi(x)
    \rangle_{\mathcal{H}}
    \\
    &=1
\end{align*}
\end{proof}

@@

\begin{table}
\caption{MAE Boston data}    
\begin{tabular}{lllll}
\toprule
    & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
\midrule
0 & 3.826326 & 2.845989 & 2.965490 & 2.810494 \\
\bottomrule
\end{tabular}

@@

\begin{table}
    \caption{MAE Abalone data}
    \begin{tabular}{lllll}
    \toprule
     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
    \midrule
    0 & 1.627627 & 1.574179 & 1.499522 & 1.498583 \\
    \bottomrule
    \end{tabular}
\end{table}

@@

\begin{table}
    \caption{MAE Vehicle data}
    \begin{tabular}{lllll}
    \toprule
     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
    \midrule
    0 & 1.117714 & 0.606971 & 0.752197 & 0.594292 \\
    \bottomrule
    \end{tabular}
\end{table}
What can be concluded from these numerical examples is that, on average kernel quantile regression yields better results than quantile forest \cite{meinshausen2006quantile} and gradient boosting machine quantile regression \cite {friedman2001greedy} in terms of the pinball loss as well as in terms of the mean absolute error.

@@

\newpage
\section{Src code}\label{src_code}
The whole code for the project is hosted on
\url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
\\
\begin{itemize}
    \item query: folder containing Scopus data and scripts to generate bibliometric survey plots in section \ref{literature_review}
    \item kqr.py: file implementing our custom kernel quantile regression
\end{itemize}

>>>>

\chapter{Appendix}
\section{Feature map normalization}\label{appendix:new_feature}
\begin{proof}
\begin{align*}
    \|
    \phi^{new}(x)\|_{\mathcal{H}}^{2} &= \left\|\frac{\phi(x)}{\|\phi(x)\|_{\mathcal{H}}}
    \right\|_{\mathcal{H}}^{2}
    \\
    &=
    \left\|
    \frac{\phi(x)}
    {\sqrt{k(x,x)}}
    \right\|_{\mathcal{H}}^{2}
    \\
    &=
    \langle
    \frac{\phi(x)}
    {\sqrt{k(x,x)}}
    ,
    \frac{\phi(x)}
    {\sqrt{k(x,x)}}
    \rangle_{\mathcal{H}}
    \\
    &=
    \frac{1}{\sqrt{k(x,x)^{2}}}
    \langle
    \phi(x)
    ,
    \phi(x)
    \rangle_{\mathcal{H}}
    \\
    &=1
\end{align*}
\end{proof}

@@

\begin{table}
\caption{Mean absolute error Boston data}    
\begin{tabular}{lllll}
\toprule
    & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
\midrule
0 & 3.826326 & 2.845989 & 2.965490 & 2.810494 \\
\bottomrule
\end{tabular}

@@

\begin{table}
    \caption{Mean absolute error Abalone data}
    \begin{tabular}{lllll}
    \toprule
     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
    \midrule
    0 & 1.627627 & 1.574179 & 1.499522 & 1.498583 \\
    \bottomrule
    \end{tabular}
\end{table}

@@

\begin{table}
    \caption{Mean absolute error Vehicle data}
    \begin{tabular}{lllll}
    \toprule
     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
    \midrule
    0 & 1.117714 & 0.606971 & 0.752197 & 0.594292 \\
    \bottomrule
    \end{tabular}
\end{table}
What can be concluded from these numerical examples is that, on average kernel quantile regression yields better results than quantile forest \cite{meinshausen2006quantile} and gradient boosting machine quantile regression \cite {friedman2001greedy} in terms of the pinball loss as well as in terms of the mean absolute error.

@@

\newpage
\section{Source code}\label{src_code}
The whole code for the project is hosted on
\url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
\\
\begin{itemize}
    \item query: folder containing Scopus data and scripts to generate bibliometric survey plots in section \ref{literature_review}
    \item kqr.py: file implementing our custom kernel quantile regression
    \item thesis: folder containing tex files
    \item experiments: folder containing train, test script for each of the experiments carried out in the thesis alongside with the binary files of the pickled models.
\end{itemize}

### 2024-05-14 thesis/input_files/data_exploratory.tex
May 13th 2024

- Explain how data has been retrieved.

- Data provider EPEX(entsoe retrieves its data)

- Explain the ETL(Extract Transform Load) pipeline
I set up.

- Explore data in order to get useful insights for
how to tune the models to get the most of them.

- correlation between temperatures and load

- Correlation and auto correlation plots

- Split train and test dataset
Explain carefully why it is important to carry
out an out of sample test and not in sample.
In sample test involves look ahead bias because we are
fitting the model on the data we want to predict,
thus it overfits on the data considered but it does
not generalize well.

This section covers the datasets used in our experiments, attributes and features will be thoroughly described.
Next, we will explain the ETL pipeline that we set up in order to ease the workflow of our comparison studies.
Finally, in order to to bettern understand patters within the data, we carry out an exploratory data analysis.

\section{Dataset}
The chosen datasets for probabilistic load forecasting, were the data from the GEFCom 2014 competition. The data is freely hosted on Dr. Hong blog \cite{hong2016probabilistic}.
The main reason was that, these dataset are considered an a excellent test case for comparing predictive models between the EF community. Additionally, the scores of the competing models are freely available, this enables us to carry out a clean and transparent comparative study.
The GEFCom 2014 consisted of four tracks: price, load, wind power and solar power forecasting. 
The GEFCom2014 folder contains a zip file for each of the four tracks.
This thesis work is focuesed on providing forecasts for the load and price quantities.
\\
\subsection{Price track}
In the price forecasting track, the goal was to predict the electricity price for the next 24 hours of a single zone. 
The data can be found in the GEFCom2014-P\_V2 zip file alongside a set of instructions and the benchmark forecasts.
This data consisted of time series for the locational marginal price, for the zonal load and for the system load. The data covers the time interval ranging from the 1st of January 2011 to the 15th of June 2013, see figure \ref{fig:price_track_fig1}. 
Additionally, as the competition went on, the real observed data of the previous tasks were made available.
The fifteen target days ordered by task number are: 16/06/2013, 17/06/2013, 24/06/2013, 04/07/2013, 09/07/2013, 13/07/2013, 16/07/2013, 18/07/2013, 19/07/2013, 20/07/2013, 24/07/2013, 25/07/2013, 07/12/2013, 08/12/2013. 
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/price_track_fig1.png}
    \caption{Price track}
    \label{fig:price_track_fig1}
\end{figure}
\subsection{Load track}
In the load track, contestant were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds. To get started, in the first round, organizers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010). 
As for the price track, the true observed data from the previous track were made available as the competition progressed.
\\
The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file. Within this subfolder, we a txt with the competition instructions and 15 folders one for each of the consecutive tasks. For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
Each data from the various task folders are shifted by one months between others.
\\
In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_V2 file contained also in the GEFCom2014 Data directory.
% Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance. This choice was motivated by the need 
% to give a proportial weight to each task when averaging between them; that is in order to not penalise too much the tasks with higher point spreads. Furthermore, additional factors such as frequency of benchmark outperformance, and robustness and clarity in methodology explaination.
\\
Notice, the pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.

\section{ETL}

\section{EDA}

>>>>

% - Explain how data has been retrieved.

% - Data provider Energycharts(entsoe retrieves its data)

% - Explain the ETL(Extract Transform Load) pipeline
% I set up.

% - Explore data in order to get useful insights for
% how to tune the models to get the most of them.

% - correlation between temperatures and load, wind speeds and load

% - Split train and test dataset
% Explain carefully why it is important to carry
% out an out of sample test and not in sample.
% In sample test involves look ahead bias because we are
% fitting the model on the data we want to predict,
% thus it overfits on the data considered but it does
% not generalize well.

This section covers the datasets used in our experiments, in addition, attributes and features will be thoroughly described.
Next, we will explain the ETL pipeline that we set up in order to ease the workflow of our comparison studies.
Finally, in order to to bettern understand patters within the data, we carry out an exploratory data analysis.

\section{GEFCom2014}
The first class of datasets we identified were the ones from the GEFCom 2014 competition. The data is freely hosted on Dr. Hong blog \cite{hong2016probabilistic}.
The main reason is that, these dataset are considered an a excellent test case for comparing predictive models between the EF community. Additionally, the scores of the competing models are freely available, this enables us to carry out a clean and transparent comparative study.
The GEFCom 2014 consisted of four tracks: price, load, wind power and solar power forecasting. 
The GEFCom2014 folder contains a zip file for each of the four tracks.
This thesis work is focuesed on providing forecasts for the load and price quantities.
\\
\subsection{GEFCom2014 Price track}
In the price forecasting track, the goal was to predict the electricity price for the next 24 hours of a single zone. 
The data can be found in the GEFCom2014-P\_V2 zip file alongside a set of instructions and the benchmark forecasts.
This data consisted of time series for the locational marginal price, for the zonal load and for the system load. The data covers the time interval ranging from the 1st of January 2011 to the 15th of June 2013, see figure \ref{fig:price_track_fig1}. 
Additionally, as the competition went on, the real observed data of the previous tasks were made available.
The fifteen target days ordered by task number are: 16/06/2013, 17/06/2013, 24/06/2013, 04/07/2013, 09/07/2013, 13/07/2013, 16/07/2013, 18/07/2013, 19/07/2013, 20/07/2013, 24/07/2013, 25/07/2013, 07/12/2013, 08/12/2013. 
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/price_track_fig1.png}
    \caption{Price track}
    \label{fig:price_track_fig1}
\end{figure}
\subsubsection{EDA}
In exploring this data we started off by plotting the time series of the zonal prices alongside with the toal load and zonal load, see figure \ref{fig:price_track_fig1}.
After that, we plotted the scatter plot for inspecting the correlation between the predictors and the dependent variable, figure \ref{fig:gefcom_zonal_price_vs_zonal_load} and figure \ref{fig:gefcom_zonal_price_vs_total_load}. From these visualisation, we can see their goodness for predicting zonal price since there exist a clear correlation between the candidate predictors and the target variable.

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/gefcom_zonal_price_vs_zonal_load.png}
    \caption{Zonal price versus zonal load}
    \label{fig:gefcom_zonal_price_vs_zonal_load}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/gefcom_zonal_price_vs_total_load.png}
    \caption{Zonal price versus total load}
    \label{fig:gefcom_zonal_price_vs_total_load}
\end{figure}

\subsection{Load track}
In the load track, contestant were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds. To get started, in the first round, organizers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010). 
As for the price track, the true observed data from the previous track were made available as the competition progressed.
\\
The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file. Within this subfolder, we have a txt with the competition instructions and 15 subfolders one for each of the consecutive tasks. For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
Each data from the various task folders are shifted by one months between others.
\\
In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_V2 file contained also in the GEFCom2014 Data directory.
% Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance. This choice was motivated by the need 
% to give a proportial weight to each task when averaging between them; that is in order to not penalise too much the tasks with higher point spreads. Furthermore, additional factors such as frequency of benchmark outperformance, and robustness and clarity in methodology explaination.
\\
Notice, the pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.
\subsubsection{EDA}
We applied the same kind of EDA to the GEFCom load data. Figure \ref{fig:gefcom_load_historical} depicts the hisorical load while figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature. Finally figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the the load and the weather.
These plots motivate the selection of weather temperature as predictor for the load.

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/gefcom_load_historical.png}
    \caption{Historical load}
    \label{fig:gefcom_load_historical}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/gefcom_w_avg_historical.png}
    \caption{Historical weather}
    \label{fig:gefcom_w_avg_historical}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/gefcom_load_vs_w_avg.png}
    \caption{Load versus weather}
    \label{fig:gefcom_load_vs_w_avg}
\end{figure}

\subsection{GEFCom ETL}
The provided data is raw, hence the need for a standardized extract transform and load pipeline. 
Firstly, we drop all Nans and zeros. Notice, the provided data for the challenge has a timestamp column not coherent to any time format. Thus, we first pass the data through a method extracting time information: hour, day, month and year. This info is then combined into a datetime object, we selected the ISO format.
Next, since the data does not provide the location for the target utility, the best we can do is selecting the average of weather temperatures as predictor variables. This independent variable is named w\_avg.
The same thing applies in the cleaning of the test data with the addition of an if clause for handling the different naming convention and formatting used in the last track of the challenge. The python etl pipeline has been then automated to loop every task folder through bash scripting.

\section{Germany Switzerland Austria data}
Next, we identified various data source for Germany, Switzerland and Austria in order to further
investigate the techniques studied in this thesis work.
Taking inspiration from the GEFCom2014 format, we firstly created a load track for Germany and Switzerland. 
The source of data has been the \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts} website. It provides data for data for electricity power, electricity potential capacity, energy generation, market prices, environmental measurements and also scenario simulations. Notice, this data is retrieved from the European network of transmission system operators for electricity \href{https://www.entsoe.eu/}{ENTSOE} and made available thanks to the \href{https://ise.fraunhofer.de/en.html}{Fraunhofer Institute for Solar Energy Systems ISE}.
\\
In this experiment, the goal is comparing kernel quantile regression to the other state of the art quantile regressors, see section \ref{ch:prob}.
The target dependent variable is the national load while as predictors we took
the weather temperature, the wind speed, the date expressed in terms of the three attributes day, month and hour, a categorical variable for the day of the week and a binary categorical variable accounting for the holiday effect.

In the second experiment we compared various kernels in the context of load foreacsting.
For this study we considered the data of \href{https://zenodo.org/records/7907883}{SECURES-Met}. This data consists of various metereological attributes relevant for modelling electricity demand. The dataset is made up of historical data until end of 2020, while from 2021 onwards the data consists of forecasts made by a metereological model developed by the European Centre for Medium-Range Weather Forecasts (ECMWF). Notice, the forecasted data is splitted in two files according to the scenarios RCP4.5 and RCP8.5. In our experiment we considered the former. 
In this setting the independent variables available are 
direct irradiation, global radiation, power from reservoir plants, power from run of the river plants (ROR), weather temperature and wind potential.

\subsubsection{ETL}
From energy charts we downloaded the historical load along with the weather temperatures and wind speeds forecasts. Notice, the data request to the environmental variables at hourly frequency for a whole year is broken. Therefore, we downloaded each month separately and then transformed into a single csv for the whole year.

For what concers the data from SECURES-Met we first identified the data interesting to us. Next, we formatted all the csv to a coherent format, gave coloumns meaningful names, sliced for the years 2020 (train) and 2021 (test) and finally we merged all single csv together.

All the manipulations involved in the ETL pipeline have been carried out by combining python and bash scripting.

\subsubsection{EDA}
Following are reported the EDA plots alongside with insights that can be extracted from them. We choose to focus the following analysis on the swiss data only given that the plots for the german data are pretty similar.
To get started we visualized the historical load data in figure \ref{fig:CH_historical_load_2021}. As expected, electricity load exhibits a seasonal patter with highs in the winter and lows during the summer season.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_historical_load_2021.png}
    \caption{Switzerland historical load}
    \label{fig:CH_historical_load_2021}
\end{figure}
Next we built scatter plots of the independent continous variables against the load as means of visualizing correlation between them.
Figure \ref{fig:CH_load_vs_temperature_2021} plots load against temperatures while figure \ref{fig:CH_load_vs_wind_speed_2021} depicts load against wind speed. From these plots we can clearly see there exist a relationship that ties wind speeds and temperatures to load.

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_load_vs_temperature_2021.png}
    \caption{Load vs temperature Switzerland 2021}
    \label{fig:CH_load_vs_temperature_2021}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_load_vs_wind_speed_2021.png}
    \caption{Load vs wind speed Switzerland 2021}
    \label{fig:CH_load_vs_wind_speed_2021}
\end{figure}

Following, we used box plots in order to visualize the relations between categorical independent variables and the dependent one. Such visualisation are reported in figure \ref{fig:CH_day_of_week_boxplot_2021} and figure \ref{fig:CH_is_holiday_boxplot_2021}. We can see that weekdays and workdays involve a higher load and posses a wider range.

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_day_of_week_boxplot_2021.png}
    \caption{Load versus day of week boxplot}
    \label{fig:CH_day_of_week_boxplot_2021}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_is_holiday_boxplot_2021.png}
    \caption{Load versus day type boxplot}
    \label{fig:CH_is_holiday_boxplot_2021}
\end{figure}

We conclude this EDA by reporting the heatmap of independent variables alongside with the load in figure \ref{fig:CH_heatmap_2021}. Temperature, wind speed and day of week show a meaningful correlation. This is a clear indication of their suitability as predictors for forecasting the load.

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_heatmap_2021.png}
    \caption{Switzerland load heatmap (2021)}
    \label{fig:CH_heatmap_2021}
\end{figure}

### 2024-05-14 thesis/input_files/implementation.tex
May 13th 2024

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of statsmodels.regression.quantile\_regression.QuantReg has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of sklearn.ensemble.GradientBoostingRegressor has been used.
\subsection{Quantile forest}
The implementation of quantile\_forest.RandomForestQuantileRegressor has been used. This estimator is compatible with scikit-learn.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented open source library, thus the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialized by providing a quantile, the regularization term C and the rbf kernel bandwidth gamma; the latter are the two hyperparameters of our custom estimator.
\\
In the fit method, we set up and solve the convex optimisation problem by using the cvxopt library. 
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the Q matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them an provide as a unique matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix X\_eval of independent variables, next we compute the kernel matrix between X\_train and X\_eval and obtain y\_eval with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules.
\\
Up to now, there are only two open source implementation of the quantile kernel regression. Nevertheless they are both in R, that is there exists no python, matlab or julia open source implementation. Following are reported the results of a comparative study between our own implementation and the one of the R library kernlab
% (i guess it is in C or C++ and then binded to R).
\subsubsection{Python versus R implementation}
In this section, a comparison study has been carried out in order to inspect the competitiveness of our implementation with the existing one for the R programming language.

>>>>

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of statsmodels.regression.quantile\_regression.QuantReg has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of sklearn.ensemble.GradientBoostingRegressor has been used.
\subsection{Quantile forest}
The implementation of quantile\_forest.RandomForestQuantileRegressor has been used. This estimator is compatible with scikit-learn.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented open source library, thus the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialized by providing a quantile, the regularization term C and the rbf kernel bandwidth gamma; the latter are the two hyperparameters of our custom estimator.
\\
In the fit method, we set up and solve the convex optimisation problem by using the cvxopt library. 
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the Q matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them an provide as a unique matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix X\_eval of independent variables, next we compute the kernel matrix between X\_train and X\_eval and obtain y\_eval with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules. Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}, functions supported by our kernel quantile regression include: gaussian rbf, laplacian, mater 1/2, matern 3/2, matern 5/2, linear, cosine, sigmoid, periodic, polynomial and custom composition of kernels.
% NOTICE, these kernel implementations differ in the form of passed parameters in the former we pass in the numerator gamma while in the latter we pass l, which stands for length scale in the denominator. To cross validate overt the same range of parameter across the different kernels we made sura that gamma=1/l.
\\
Up to now, there are only two open source implementation of the quantile kernel regression. Nevertheless they are both in R, that is there exists no python, matlab or julia open source implementation. Following are reported the results of a comparative study between our own implementation and the one of the R library kernlab
% (i guess it is in C or C++ and then binded to R).
% \subsubsection{Python versus R implementation}
% In this section, a comparison study has been carried out in order to inspect the competitiveness of our implementation with the existing one for the R programming language.
